---
title: "combined_km"
author: "Marietta_Dalamanga"
date: "2024-07-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Survival analysis, Kaplan-Meier curves

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(survminer)
library(survival)
library(data.table)
library(coin)

```

The data we will use is the dataset with the most common conditions (conditions found on more than 2000 animals) and breeds(8 breeds account for the 91.5% of all animals)
```{r}

survival_data <- read.csv("./data/common_conditions_breed_data.csv")
```

```{r}
colnames(survival_data)
```

All categorical variables as factor
```{r}
survival_data[,c("Herd_Id", "Abbatoir", "Dvo_Code", "Breed", "Sex", "Abattoir_._Died_on_Farm", "Beef_or_Dairy")] = 
  lapply(survival_data[,c("Herd_Id", "Abbatoir", "Dvo_Code", "Breed", "Sex", "Abattoir_._Died_on_Farm", "Beef_or_Dairy")],
                                function(x) as.factor(x))
```

```{r}
healthy_data <- survival_data %>%
  filter(Total_Conditions_per_Animal == 0)
unhealthy_data <- survival_data %>%
  filter((Total_Conditions_per_Animal >0))
```

We will perform Kaplan-Meier analysis on days alive

### Baseline survival curve

```{r}
survival_object <- Surv(survival_data$days_alive, survival_data$status)
healthy_object <- Surv(healthy_data$days_alive, healthy_data$status)
unhealthy_object <- Surv(unhealthy_data$days_alive, unhealthy_data$status)
```


```{r}
# function to get the survival on different time points
km_survival_time <- function(model_fit) {
  survival_time  <- data.frame(time = model_fit$time,
                                  n.risk = model_fit$n.risk,
                                  n.event = model_fit$n.event,
                                  n.censor = model_fit$n.censor,
                                  surv = model_fit$surv,
                                  upper = model_fit$upper,
                                  lower = model_fit$lower)
  return(survival_time)
}

```


```{r}
# function for plot
curve <- function(model, var){
  km_curve <- ggsurvplot(model,
          pval = TRUE, 
          conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          surv.median.line = "hv", # Specify median survival
          
          ggtheme = theme_bw()# Change ggplot2 theme
          )+
  labs(title=paste("Survival per", var, "in days"), x="Days", y="Survival_probability") 
  return(km_curve)
}

```

Overall survival in days
```{r}
baseline_km_all_data <- survfit(survival_object ~ 1, data = survival_data)
summary(baseline_km_all_data)$table
baseline_km_all_data_months <- survfit(Surv(months_alive, status) ~ 1, data=survival_data)
```

```{r}
km_survival_time(baseline_km_all_data)
km_survival_time(baseline_km_all_data_months)
km_survival_time(survfit(Surv(years_alive, status) ~1, data=survival_data))
```


```{r}
km_baseline_days <- ggsurvplot(baseline_km_all_data,
          pval = FALSE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "blue")+
  labs(title="Baseline survival in days", x="Days", y="Survival_probability") 
km_baseline_days
ggsave("./data/combined_figs/baseline_km_all_data.png", plot=baseline_km_all_data$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_of_baseline_km_all_data.png", plot=baseline_km_all_data$table)
```

```{r}
km_baseline_months <- ggsurvplot(baseline_km_all_data_months,
          pval = FALSE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "blue")+
  labs(title="Baseline survival in months", x="Days", y="Survival_probability") 
km_baseline_months
km_baseline_years <- ggsurvplot(survfit(Surv(years_alive, status) ~1, data=survival_data),
          pval = FALSE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "blue")+
  labs(title="Baseline survival in years", x="Days", y="Survival_probability") 
km_baseline_years
```

baseline survival for healthy and unhealthy animals

```{r}
baseline_km_healthy <- survfit(healthy_object ~ 1, data = healthy_data)
baseline_km_unhealthy <- survfit(unhealthy_object ~ 1, data = unhealthy_data)
summary(baseline_km_healthy)$table
summary(baseline_km_unhealthy)$table
```

```{r}
# all curves
fit <- list(all = baseline_km_all_data, healthy = baseline_km_healthy, unhealthy=baseline_km_unhealthy)
all_baseline_curves <- ggsurvplot(fit, data = survival_data, combine = TRUE, # Combine curves
           risk.table = TRUE,                  # Add risk table
           conf.int = TRUE,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           )
ggsave("./data/combined_figs/all_baseline_curves.png", plot=all_baseline_curves$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_of_all_baseline_curves.png", plot=all_baseline_curves$table)
all_baseline_curves
```

survival curve on healthy and unhealthy animals based on total conditions
```{r}
dat_all <- survival_data
dat_all$Total_Conditions_per_Animal[dat_all$Total_Conditions_per_Animal >0] <- 'unhealthy'
dat_all$Total_Conditions_per_Animal[dat_all$Total_Conditions_per_Animal == 0] <- 'healthy'
colnames(dat_all)
dat_all <- rename(dat_all, Health_status= Total_Conditions_per_Animal)
healthy_vs_unhealthy <- survfit(survival_object ~ Health_status, data = dat_all)
summary(healthy_vs_unhealthy)$table
```



```{r}
curve_healthy_vs_unhealthy <- ggsurvplot(healthy_vs_unhealthy,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Health_status", # Change risk table color by groups
          linetype = "Health_status", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()
          )+
    labs(title="Survival in days by health status", x="Days", y="Survival_probability") 
curve_healthy_vs_unhealthy
ggsave("./data/combined_figs/curve_healthy_vs_unhealthy.png", plot=curve_healthy_vs_unhealthy$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_curve_healthy_vs_unhealthy.png", plot=curve_healthy_vs_unhealthy$table)
```


```{r}
fit1 <- list(all = baseline_km_all_data, healthy_vs_unhealthy)
all_baseline_curves1 <- ggsurvplot(fit, data = survival_data, combine = TRUE, # Combine curves
           risk.table = TRUE,                  # Add risk table
           conf.int = TRUE,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           )

all_baseline_curves1
```



### survival curve for sex

```{r}

sex_fit <- survfit(survival_object ~ Sex, data=survival_data)
summary(sex_fit)$table

```

```{r}
sex_fit_healthy <- survfit(healthy_object ~ Sex, data=healthy_data)
summary(sex_fit_healthy)$table
sex_fit_unhealthy <- survfit(unhealthy_object ~ Sex, data=unhealthy_data)
summary(sex_fit_unhealthy)$table
```


```{r}
km_sex_table <- data.frame(Sex = c("male", "female", "bull"), 
                           overall_median = c(summary(sex_fit)$table[1, 7], summary(sex_fit)$table[2, 7], summary(sex_fit)$table[3, 7]),
                           healthy_median = c(summary(sex_fit_healthy)$table[1, 7], summary(sex_fit_healthy)$table[2, 7], summary(sex_fit_healthy)$table[3, 7]),
                           unhealthy_median = c(summary(sex_fit_unhealthy)$table[1, 7], summary(sex_fit_unhealthy)$table[2, 7], summary(sex_fit_unhealthy)$table[3, 7]))

km_sex_table
```



```{r}
km_sex <- ggsurvplot(sex_fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Sex", # Change risk table color by groups
          #linetype = "Sex", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
         )+
    labs(title="Survival in years vs sex", x="Days", y="Survival_probability") 

ggsave("./data/combined_figs/km_sex.png", plot=km_sex$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_km_sex.png", plot=km_sex$table)
km_sex
```

Log-rank tests
```{r}
sex_diff <- survdiff(survival_object ~ Sex, data=survival_data)
sex_diff
```


```{r}
km_sex_pairwise_F_M <- survdiff(survival_object ~ Sex, data=survival_data, subset = Sex!="B")
km_sex_pairwise_F_M
km_sex_pairwise_B_M <- survdiff(survival_object ~ Sex, data=survival_data, subset = Sex!="F")
km_sex_pairwise_B_M
km_sex_pairwise_B_F <- survdiff(survival_object ~ Sex, data=survival_data, subset = Sex!="M")
km_sex_pairwise_B_F
```

```{r}
km_sex_helthy_versus_unhealthy <- survfit(survival_object ~  Health_status + Sex, data=dat_all)

summary(km_sex_helthy_versus_unhealthy)$table


km_sex_helthy_versus_unhealthy <- ggsurvplot(km_sex_helthy_versus_unhealthy,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Sex", # Change risk table color by groups
          linetype = "Health_status", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
         )+
    labs(title="Survival in days vs sex", x="Days", y="Survival_probability") 

ggsave("./data/combined_figs/km_sex_helthy_versus_unhealthy.png", plot=km_sex$plot, width = 8, height = 6)
ggsave("./data/combined_figs/km_sex_helthy_versus_unhealthy.png", plot=km_sex$table)
km_sex_helthy_versus_unhealthy
```


### survival curve for place of death

```{r}

colnames(survival_data)
survival_data <- rename(survival_data, Place_of_death=Abattoir_._Died_on_Farm) 
dat_all <- rename(dat_all, Place_of_death=Abattoir_._Died_on_Farm)
```

```{r}

place_of_death_fit_days <- survfit(survival_object ~ Place_of_death, data=survival_data)
summary(place_of_death_fit_days)$table

place_of_death_helthy_versus_unhealthy <- survfit(survival_object ~  Health_status + Place_of_death, data=dat_all)
summary(place_of_death_helthy_versus_unhealthy)$table

```


```{r}
km_place_of_death_days <- ggsurvplot(place_of_death_fit_days,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table between 318 and 326
          risk.table.col = "Place_of_death", # Change risk table color by groups
          # linetype = "Place_of_death", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"))+
    labs(title="Survival in days vs place of death", x="Days", y="Survival probability") 

ggsave("./data/combined_figs/km_place_of_death_days.png", plot=km_place_of_death_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_km_place_of_death_days.png", plot=km_place_of_death_days$table)
km_place_of_death_days

```

```{r}

km_place_of_death_days_vs_health<- ggsurvplot(place_of_death_helthy_versus_unhealthy,
          pval = FALSE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          # risk.table.col = "Place_of_death", # Change risk table color by groups
          linetype = "Health_status", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), 
          legend.labs=c("healthy, abatttoir", "healthy, farm", "unhealthy, abattoir", "unhealthy, farm")
          )+
    labs(title="Survival in days vs place of death and health status", x="Days", y="Survival_probability") 

km_place_of_death_days_vs_health 


ggsave("./data/combined_figs/km_place_of_death_days_vs_health.png", plot=km_place_of_death_days_vs_health$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_tablekm_place_of_death_days_vs_health.png", plot=km_place_of_death_days_vs_health$table)



```

```{r}
place_of_death_pairwise_unhealthy <- survdiff(survival_object ~ Place_of_death, data=dat_all, subset = Health_status!="healthy")
place_of_death_pairwise_unhealthy
place_of_death_pairwise_healthy <- survdiff(survival_object ~ Place_of_death, data=dat_all, subset = Health_status!="unhealthy")
place_of_death_pairwise_healthy

```



Type of production


days
```{r}

production_fit_days <- survfit(survival_object ~ Beef_or_Dairy, data=survival_data)

summary(production_fit_days)$table

```


```{r}
prod_fit_months <- survfit(Surv(survival_data$months_alive, survival_data$status)~Beef_or_Dairy, data=survival_data)
time_prod_months <- data.frame(time = prod_fit_months$time,
                  n.risk = prod_fit_months$n.risk,
                  n.event = prod_fit_months$n.event,
                  n.censor = prod_fit_months$n.censor,
                  surv = prod_fit_months$surv,
                  upper = prod_fit_months$upper,
                  lower = prod_fit_months$lower
                  )
time_prod_months

```


```{r}
km_production_days <- ggsurvplot(production_fit_days,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Beef_or_Dairy", # Change risk table color by groups
          linetype = "Beef_or_Dairy", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF")) +
        labs(title="Survival in years vs type of production", x="Years", y="Survival probability") 

ggsave("./data/combined_figs/km_production_days.png", plot=km_production_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_km_production_days.png", plot=km_production_days$table)
km_production_days
```

production vs health

```{r}
prod_helthy_vs_unhealthy <- survfit(survival_object ~  Health_status + Beef_or_Dairy, data=dat_all)
summary(prod_helthy_vs_unhealthy)$table
```

```{r}
km_production_vs_health <- ggsurvplot(prod_helthy_vs_unhealthy,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Beef_or_Dairy", # Change risk table color by groups
          linetype = "Health_status", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          ) +
        labs(title="Survival in years vs type of production", x="Years", y="Survival probability") 

ggsave("./data/combined_figs/km_production_vs_health.png", plot=km_production_vs_health$plot, width = 8, height = 6)
ggsave("./data/combined_figs/risk_table_km_production_vs_health.png", plot=km_production_vs_health$table)
km_production_vs_health
```

```{r}
prod_pairwise_unhealthy <- survdiff(survival_object ~ Beef_or_Dairy, data=dat_all, subset = Health_status!="healthy")
prod_pairwise_unhealthy
prod_pairwise_healthy <- survdiff(survival_object ~ Beef_or_Dairy, data=dat_all, subset = Health_status!="unhealthy")
prod_pairwise_healthy

```


Dvo Code
```{r}

DVO_fit_days <- survfit(Surv(days_alive, status) ~ Dvo_Code, data=survival_data)
summary(DVO_fit_days)$table
DVO_fit_health <- survfit(Surv(days_alive, status) ~ Dvo_Code + Health_status, data=dat_all)
summary(DVO_fit_health)$table
```


```{r}
 km_DVO_days <- ggsurvplot(DVO_fit_days,
                           data = survival_data,
                           pval = TRUE,
                           conf.int = TRUE,
                           risk.table = TRUE,
                           title = paste("Survival Curve for all DVO codes, (Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability",
    legend.labs = c(levels(survival_data$Dvo_Code))
  )
ggsave("./data/combined_figs/km_DVO_days.png", plot=km_DVO_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/km_DVO_days.png", plot=km_DVO_days$table)
km_DVO_days
```

```{r}
dvo_all <- survival_data
dvo_all$Dvo_Code <- as.character(dvo_all$Dvo_Code)
dvo <- levels(survival_data$Dvo_Code)

for(i in 1:length(dvo)){
  dvo_renamed <- dvo_all
  dvo_renamed$Dvo_Code[dvo_renamed$Dvo_Code==dvo[i]] <- dvo[i]
  dvo_renamed$Dvo_Code[dvo_renamed$Dvo_Code != dvo[i]] <- "other"
  dvo_vs_other <- survfit(survival_object ~ Dvo_Code, data = dvo_renamed)
  print(summary(dvo_vs_other)$table)
  
  plot_days <- ggsurvplot(
    dvo_vs_other,
    data = dvo_renamed,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = TRUE,
    title = paste("Survival Curve for", dvo[i], "(Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability"
  )
  
  print(plot_days)
  ggsave(paste("./data/combined_figs/byDvo/",dvo[i],".png", sep = ""), plot=plot_days$plot, width = 8, height = 6)
  
  dvo_renamed <- dvo_all
  }


```


Breed
```{r}

breed_fit_days <- survfit(Surv(days_alive, status) ~ Breed, data=survival_data)
summary(breed_fit_days)$table
breed_fit_health <- survfit(Surv(days_alive, status) ~ Breed + Health_status, data=dat_all)
summary(breed_fit_health)$table
```


```{r}
 km_breed_days <- ggsurvplot(breed_fit_days,
                           data = survival_data,
                           pval = TRUE,
                           conf.int = TRUE,
                           risk.table = TRUE,
                           title = paste("Survival Curve for all breeds, (Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability",
    legend.labs = c(levels(survival_data$Breed))
  )
ggsave("./data/combined_figs/km_breed_days.png", plot=km_breed_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/km_breed_days.png", plot=km_breed_days$table)
km_breed_days
```

```{r}
breed_all <- survival_data
breed_all$Breed <- as.character(breed_all$Breed)
breeds <- levels(survival_data$Breed)

for(i in 1:length(breeds)){
  breeds_renamed <- breed_all
  breeds_renamed$Breed[breeds_renamed$Breed==breeds[i]] <- breeds[i]
  breeds_renamed$Breed[breeds_renamed$Breed !=breeds[i]] <- "other"
  breed_vs_other <- survfit(survival_object ~ Breed, data = breeds_renamed)
  print(summary(breed_vs_other)$table)
  
  plot_days <- ggsurvplot(
    breed_vs_other,
    data = breeds_renamed,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = TRUE,
    title = paste("Survival Curve for", breeds[i], "(Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability"
  )
  
  print(plot_days)
  ggsave(paste("./data/combined_figs/byBreedNew/",breeds[i],".png", sep = ""), plot=plot_days$plot, width = 8, height = 6)
  
  breeds_renamed <- breed_all
  }


```

Abattoir
```{r}

Abattoir_fit_days <- survfit(Surv(days_alive, status) ~ Abbatoir, data=survival_data)
summary(Abattoir_fit_days)$table
Abattoir_fit_health <- survfit(Surv(days_alive, status) ~ Abbatoir + Health_status, data=dat_all)
summary(Abattoir_fit_health)$table
```


```{r}
 km_abatoirs_days <- ggsurvplot(Abattoir_fit_days,
                           data = survival_data,
                           pval = TRUE,
                           conf.int = TRUE,
                           risk.table = TRUE,
                           title = "Survival Curve for all abattoirs, (Days)",
    xlab = "Time (Days)",
    ylab = "Survival Probability",
    legend.labs = c(levels(survival_data$Abbatoir))
  )
ggsave("./data/combined_figs/km_abattoir_days.png", plot=km_abatoirs_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/km_abattoir_days.png", plot=km_abatoirs_days$table)
km_abatoirs_days
```

```{r}
abattoir_all <- survival_data
abattoir_all$Abbatoir <- as.character(abattoir_all$Abbatoir)
abattoirs <- levels(survival_data$Abbatoir)

for(i in 1:length(abattoirs)){
  abattoir_renamed <- abattoir_all
  abattoir_renamed$Abbatoir[abattoir_renamed$Abbatoir==abattoirs[i]] <- abattoirs[i]
  abattoir_renamed$Abbatoir[abattoir_renamed$Abbatoir !=abattoirs[i]] <- "other"
  abattoir_vs_other <- survfit(survival_object ~ Abbatoir, data = abattoir_renamed)
  print(summary(abattoir_vs_other)$table)
  
  plot_days <- ggsurvplot(
    abattoir_vs_other,
    data = abattoir_renamed,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = TRUE,
    title = paste("Survival Curve for", abattoirs[i], "(Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability"
  )
  
  print(plot_days)
  ggsave(paste("./data/combined_figs/byAbattoir/",abattoirs[i],".png", sep = ""), plot=plot_days$plot, width = 8, height = 6)
  
  abattoir_renamed <- abattoir_all
  }


```




years
```{r}


subset_of_conditions_1_2_3 <- survival_data %>%
  subset(survival_data$Total_Conditions_per_Animal==1 |survival_data$Total_Conditions_per_Animal==2 | survival_data$Total_Conditions_per_Animal==3)

subset_of_conditions_4_5_6 <- survival_data %>%
  subset(survival_data$Total_Conditions_per_Animal==4 |survival_data$Total_Conditions_per_Animal==5 | survival_data$Total_Conditions_per_Animal==6)

subset_of_conditions_7_8_9 <- survival_data %>%
  subset(survival_data$Total_Conditions_per_Animal==7 |survival_data$Total_Conditions_per_Animal==8 | survival_data$Total_Conditions_per_Animal==9)

total_conditions_fit_years <- survfit(Surv(years_alive, status) ~ Total_Conditions_per_Animal, data=survival_data)
subset_conditions_fit_years_1_2_3<- survfit(Surv(years_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_1_2_3)
subset_conditions_fit_years_4_5_6<- survfit(Surv(years_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_4_5_6)
subset_conditions_fit_years_7_8_9<- survfit(Surv(years_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_7_8_9)
```


```{r}

summary(total_conditions_fit_years)$table

```

```{r}
survival_total_conditions_years <- data.frame(time = total_conditions_fit_years$time,
                  n.risk = total_conditions_fit_years$n.risk,
                  n.event = total_conditions_fit_years$n.event,
                  n.censor = total_conditions_fit_years$n.censor,
                  surv = total_conditions_fit_years$surv,
                  upper = total_conditions_fit_years$upper,
                  lower = total_conditions_fit_years$lower
                  )
survival_total_conditions_years
```


```{r}
subset_1_total_conditions_years <- ggsurvplot(subset_conditions_fit_years_1_2_3, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Years", y="Survival probability") 
ggsave("./data/combined_figs/subset_1_total_conditions_years.png", plot=subset_1_total_conditions_years$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_1_total_conditions_years.png", plot=subset_1_total_conditions_years$table)
subset_1_total_conditions_years
```
```{r}
subset_2_total_conditions_years <- ggsurvplot(subset_conditions_fit_years_4_5_6, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Years", y="Survival probability") 
ggsave("./data/combined_figs/subset_2_total_conditions_years.png", plot=subset_2_total_conditions_years$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_2_total_conditions_years.png", plot=subset_2_total_conditions_years$table)
subset_2_total_conditions_years
```

```{r}
subset_3_total_conditions_years <- ggsurvplot(subset_conditions_fit_years_7_8_9, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Years", y="Survival probability") 
ggsave("./data/combined_figs/subset_3_total_conditions_years.png", plot=subset_3_total_conditions_years$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_3_total_conditions_years.png", plot=subset_3_total_conditions_years$table)
subset_3_total_conditions_years
```



days
```{r}

total_conditions_fit_days <- survfit(Surv(days_alive, status) ~ Total_Conditions_per_Animal, data=survival_data)
                     
```


```{r}

summary(total_conditions_fit_days)$table

```

```{r}
survival_total_conditions_days <- data.frame(time = total_conditions_fit_days$time,
                  n.risk = total_conditions_fit_days$n.risk,
                  n.event = total_conditions_fit_days$n.event,
                  n.censor = total_conditions_fit_days$n.censor,
                  surv = total_conditions_fit_days$surv,
                  upper = total_conditions_fit_days$upper,
                  lower = total_conditions_fit_days$lower
                  )
survival_total_conditions_days
```


```{r}
km_total_conditions_days <- ggsurvplot(total_conditions_fit_days,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in days vs total post-mortem conditions", x="Days", y="Survival probability") 
ggsave("./data/combined_figs/km_total_conditions_days.png", plot=km_total_conditions_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/km_total_conditions_days.png", plot=km_total_conditions_days$table)
km_total_conditions_days
```
```{r}
subset_conditions_fit_days_1_2_3 <- survfit(Surv(days_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_1_2_3)
subset_conditions_fit_days_4_5_6 <- survfit(Surv(days_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_4_5_6)
subset_conditions_fit_days_7_8_9 <- survfit(Surv(days_alive, status) ~ Total_Conditions_per_Animal, data=subset_of_conditions_7_8_9)
```

```{r}
subset_1_total_conditions_days <- ggsurvplot(subset_conditions_fit_days_1_2_3, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Years", y="Survival probability") 
ggsave("./data/combined_figs/subset_1_total_conditions_days.png", plot=subset_1_total_conditions_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_1_total_conditions_days.png", plot=subset_1_total_conditions_days$table)
subset_1_total_conditions_days
```

```{r}
subset_2_total_conditions_days <- ggsurvplot(subset_conditions_fit_days_4_5_6, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Days", y="Survival probability") 
ggsave("./data/combined_figs/subset_2_total_conditions_days.png", plot=subset_2_total_conditions_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_2_total_conditions_days.png", plot=subset_2_total_conditions_days$table)
subset_2_total_conditions_days
```


```{r}
subset_3_total_conditions_days <- ggsurvplot(subset_conditions_fit_days_7_8_9, 
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.y.text = FALSE,
          risk.table.col = "Total_Conditions_per_Animal", # Change risk table color by groups
          linetype = "Total_Conditions_per_Animal", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
        ) +
  labs(title="Survival in years vs total post-mortem conditions", x="Days", y="Survival probability") 
ggsave("./data/combined_figs/subset_3_total_conditions_days.png", plot=subset_3_total_conditions_days$plot, width = 8, height = 6)
ggsave("./data/combined_figs/subset_3_total_conditions_days.png", plot=subset_3_total_conditions_days$table)
subset_3_total_conditions_days
```







Conditions
```{r}
condition_columns <- colnames(survival_data)[21:44]
```


```{r}

surv_obj_days <- Surv(time = survival_data$days_alive, event = survival_data$status)
surv_obj_years <- Surv(time = survival_data$years_alive, event = survival_data$status)

```


```{r}
# Loop through the condition columns
for (i in 1:length(condition_columns)) {
  # Fit survival models for the condition
  condition_name <- condition_columns[i]
  fit_days <- survfit(surv_obj_days ~ survival_data[[condition_name]])
  fit_years <- survfit(surv_obj_years ~ survival_data[[condition_name]])
  
  # Create plots for days
  plot_days <- ggsurvplot(
    fit_days,
    data = survival_data,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = FALSE,
    title = paste("Survival Curve for", condition_name, "(Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability"
  )
  
  print(plot_days)
  ggsave(paste("./data/combined_figs/byconditionNew/",condition_name,".png", sep=""), plot=plot_days$plot, width = 8, height = 6)
}

```


One hot encoding for breed
```{r}
# Create a column for each unique breed
breed_wide_data <- survival_data %>%
  mutate(BreedPresent = 1) %>%
  pivot_wider(names_from = Breed, values_from = BreedPresent, values_fill = list(ConditionPresent = 0))
```


```{r}
breed_wide_data[is.na(breed_wide_data)] <- 0
breed_wide_data
```

```{r}
surv_obj_days <- Surv(time = breed_wide_data$days_alive, event = breed_wide_data$status)
surv_obj_years <- Surv(time = breed_wide_data$years_alive, event = breed_wide_data$status)
```

Breeds "GRB" & "LID" doesn't exist in the combined with weather dataset
```{r}
colnames(breed_wide_data[, 93:184])
colSums(breed_wide_data[, 93:184])
breed_wide_data$SI
```



```{r}
breed_columns <- c("HOL", "FR", "LIM", "AA", "HER", "BB", "CH","MB","DAQ","SHB","SIM" ,"FKV", "SH","AYR" ,"SAL" ,"SBR","SWR", "JER" ,"DEX" ,"LH",
                   "NR","BGA" ,"SHD","IM","LNR","PAR","ROT","GAL","LUI","ST","MRI","DR","HI","AU","MAJ","PIE","WP","KE" ,"EA","SD","ROM","BW","STL",
                   "CHI","BI", "BAZ","SPK","BAR","GU","BDM","GE","UNK","WAG","WG","CHL","WB","BU","TYB","AR","BBR","GLO","ALL","BRA","DEV","JEX",
                   "CAN","BAL","BRB","LBW","BLG","LAK", "MAR","ENP","FRE","BMS","NOR","BAN","SU","RPL", "MG","BEE","ARM","MAL","VAN","DN",
                   "VA","PUS","PIN","ANK","BRH","AUB","SI")
length(breed_columns)
```

Plot survival curves for each breed
```{r}

# Loop through the condition columns
for (i in 1:92) {
  # Fit survival models for the condition
  breed_name <- breed_columns[i]
  fit_days <- survfit(surv_obj_days ~ breed_wide_data[[breed_name]])
  fit_years <- survfit(surv_obj_years ~ breed_wide_data[[breed_name]])
  
  # Create plots for days
  plot_days <- ggsurvplot(
    fit_days,
    data = breed_wide_data,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = FALSE,
    title = paste("Survival Curve for", breed_name, "(Days)"),
    xlab = "Time (Days)",
    ylab = "Survival Probability"
  )
  
  print(plot_days)
  ggsave(paste("./data/combined_figs/bybreed/",breed_name,".png"), plot=plot_days$plot, width = 8, height = 6)
}
```


Plot the curves for top 6 breeds

```{r}
N <- 6
top_6_breeds <- survival_data %>%
  count(Breed, sort=TRUE) %>%
  top_n(N, n) %>%
  pull(Breed)


filtered_top_6_breeds <- survival_data %>%
  filter(Breed %in% top_6_breeds)


```


Top 8 breeds years
```{r}

Top_6_breeds_fit_years <- survfit(Surv(years_alive, status) ~ Breed, data=filtered_top_6_breeds)
                     
```


```{r}

summary(Top_6_breeds_fit_years)$table

```

```{r}
survival_top_6_breeds_years <- data.frame(time = Top_6_breeds_fit_years$time,
                  n.risk = Top_6_breeds_fit_years$n.risk,
                  n.event = Top_6_breeds_fit_years$n.event,
                  n.censor = Top_6_breeds_fit_years$n.censor,
                  surv = Top_6_breeds_fit_years$surv,
                  upper = Top_6_breeds_fit_years$upper,
                  lower = Top_6_breeds_fit_years$lower
                  )
survival_top_6_breeds_years
```


```{r}
ggsurvplot(Top_6_breeds_fit_years,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Breed", # Change risk table color by groups
          linetype = "Breed", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
)
```


Breed by production type survival curve
Dairy:
```{r}
top_dairy_breeds <- filtered_top_6_breeds %>%
  filter(Beef_or_Dairy=="Dairy")


top_dairy_breeds_fit_years <- survfit(Surv(years_alive, status) ~ Breed, data=top_dairy_breeds)

km_top_dairy_breed_years <- ggsurvplot(top_dairy_breeds_fit_years,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Breed", # Change risk table color by groups
          linetype = "Breed", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
)
km_top_dairy_breed_years
ggsave("./data/combined_figs/bybreed/km_top_dairy_breed_years.png", plot=km_top_dairy_breed_years$plot, width = 8, height = 6)
```

Breed by production type survival curve
Beef:
```{r}
top_beef_breeds <- filtered_top_6_breeds %>%
  filter(Beef_or_Dairy=="Beef")

top_beef_breeds_fit_years <- survfit(Surv(years_alive, status) ~ Breed, data=top_beef_breeds)

km_top_6_beef_breeds_years <- ggsurvplot(top_beef_breeds_fit_years,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "Breed", # Change risk table color by groups
          linetype = "Breed", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
)
km_top_6_beef_breeds_years
ggsave("./data/combined_figs/bybreed/km_top_6_beef_breeds_years.png", plot=km_top_6_beef_breeds_years$plot, width = 8, height = 6)
```

Save dataset breed wide dataset in a new csv file

```{r}
write.csv(breed_wide_data, file="./data/breed_wide.csv")
```


