---
title: "fit_distributions"
author: "Marietta_Dalamanga"
date: "2024-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(survival)
library(survminer)
library(ggplot2)
library(ggpubr)
library(fitdistrplus)
library(logspline)
library(fitur)
library(actuar)
```



```{r}
data <- read.csv("./data/common_conditions_breed_data.csv")
excluded_days <- data$days_alive <=1
data_excluded_days <- data[!excluded_days, ]
summary(data_excluded_days$days_alive)
```
```{r}
data_excluded_days[,c("Herd_Id", "Abbatoir", "Dvo_Code", "Breed", "Sex", "Abattoir_._Died_on_Farm", "Beef_or_Dairy")] = 
  lapply(data_excluded_days[,c("Herd_Id", "Abbatoir", "Dvo_Code", "Breed", "Sex", "Abattoir_._Died_on_Farm", "Beef_or_Dairy")],
                                function(x) as.factor(x))
healthy_animals <- data_excluded_days %>%
  filter(Total_Conditions_per_Animal==0)
healthy_animals <- healthy_animals[, c(14, 15, 3, 6, 7, 8, 11, 20)]
healthy_animals
```

create the dataset all animals without the conditions columns include this time and the Total_Conditions_per_Animal column
```{r}
colnames(data_excluded_days)
all_animals_no_conditions <- data_excluded_days[, c(14, 15, 3, 6, 7, 8, 11, 13, 20)]
```



```{r}
# Here we construct all possible formulas using these variables only...
predictors <- colnames(healthy_animals[3:8])

#' Formula with combination of k variables
#' 
#' @param predictors list of variables (strings)
#' @param k tuple size
#' @returns list of all possible formulas with k variables
formulas_of_len <- function(predictors, k) {
  mdl_prefix <- "Surv(days_alive, status) ~"
  all_formulas_len_k <- c()
  combs <- combn(predictors, k, simplify = FALSE)
  num_combs <- length(combs)
  for (j in 1:num_combs) {
    comb_j <- combs[j][[1]]
    comb_j_joined <- paste(comb_j, collapse = "+")
    comb_j_joined <- paste(mdl_prefix, comb_j_joined)
    all_formulas_len_k <- c(all_formulas_len_k, comb_j_joined)
  }
  return(all_formulas_len_k)
}

#' Wrapper function that returns all possible formulas (list of strings)
#' There are 2^N - 1 formulas, where N is the number of predictors
#' 
#' @param predictors list of variables (strings)
#' @returns all possible formulas (list of strings)
all_possible_formulas <- function(predictors) {
  all_formulas <- c()
  num_predictors <- length(predictors)
  for (i in 1:num_predictors) {
      these_combinations <- formulas_of_len(predictors, i)
      all_formulas <- c(all_formulas, these_combinations)
  }
  return(all_formulas)
}
```


```{r}
#' Exhaustive AFT training for all possible combinations of predictors
#' 
#' @param training_data data frame of training data (e.g., healthy_animals, or
#'                      all_animals_no_conditions)
#' @param aft_distributions list of parametric distributions (list of str)
#' @param predictors list of predictors (variables) (list of str)
#' @param destination_folder the metadata (AIC, BIC, etc) and the the models 
#'                      will be stored in RDS files in models/{destination_folder}/
#'
exhaustive_survreg_modelling <- function(training_data, aft_distributions, predictors, destination_folder) {
    # Generate formulas for all combinations of predictors
    formulas <- all_possible_formulas(predictors)
    # Initialize lists to store AIC, loglik, and BIC for each model
    aic <- c()
    bic <- c()
    llik <- c()
    mdl_name <- c()
    mdl <- c()
    num_formulas <- length(formulas)
    num_distributions <- length(aft_distributions)
    num_combinations <- num_formulas * num_distributions
    i <- 1
    # Loop over each AFT distribution and each formula, fit the model, and store AIC, loglik, and BIC
    for (aft_dist in aft_distributions) {
      for (formula in formulas) {
        # Fit the AFT model
        tryCatch({
          surv_model <- survreg(as.formula(formula), data = training_data, dist = aft_dist)
          # Calculate AIC, log-likelihood, and BIC
          aic_value <- as.numeric(AIC(surv_model))
          bic_value <- as.numeric(BIC(surv_model))
          llik_value <- as.numeric(logLik(surv_model))
          model_name <- gsub(" ", "", paste("AFT", aft_dist, formula, sep = "::"), fixed = TRUE)
          print("---------------------------------------------")
          print(paste("MODEL ", i, " / ", num_combinations, "  (", 100*i/num_combinations, "%)"))
          print(paste("Formula    : ", formula))
          print(paste("Model name : ", model_name))
          print(paste("AIC        : ", aic_value))
          print(paste("ell        : ", llik_value))
          print(paste("BIC        : ", bic_value))
          aic <- c(aic, aic_value)
          bic <- c(bic, bic_value)
          llik <- c(llik, llik_value)
          mdl_name <- c(mdl_name, model_name)
          mdl <- c(mdl, surv_model)
          print(".............................................")
          
          if (i %% 5 == 0) {
            print("Saving temporary data...")
            saveRDS(data.frame(aic, bic, llik, mdl_name), file="temp/temp_data.rds")  
            saveRDS(mdl,  file="temp/modelz.rds")  
            print("Data saved successful in temp/temp_data.rds and models in temp/modelz.rds :-)")
          }
          
          i <- i + 1
          
        }, error = function(e) {
          print(e)
          print(paste("Error happened while trying to deal with dist: ", aft_dist, " for formula ", formula))
          i <- i + 1
        })
      }
    }
    print("Modelling complete - saving final data in destination folder")
    meta_fname <- paste("models", destination_folder, "meta.rds", sep="/")
    models_fname <- paste("models", destination_folder, "models.rds", sep="/")
    saveRDS(data.frame(aic, bic, llik, mdl_name), file=meta_fname)  
    saveRDS(mdl,  file=models_fname)  
}
```


```{r}
# Here we construct all possible formulas using these variables only...
predictors <- colnames(all_animals_no_conditions[3:9])
# destination folder for final model objects
destination_folder <- "all_data"
# Define the AFT distributions you want to use
aft_distributions <- c("exponential", "weibull", "lognormal", "loglogistic")
exhaustive_survreg_modelling(all_animals_no_conditions, aft_distributions, predictors, destination_folder)
```


```{r}
colnames(data_excluded_days)
```

```{r}
destination_folder <- "test"

fname
```


all animals no conditions
```{r}
mdl_name
combinatorial_modelling_results_all <- data.frame(mdl_name, aic, bic, llik)
```

```{r}
aft_model <-survreg()

# Convert model_results into a dataframe
aft_model_results_df <- data.frame(
  Model = names(aft_comb_results),
  AIC = sapply(aft_comb_results, function(x) x$AIC),
  LogLikelihood = sapply(aft_comb_results, function(x) x$LogLikelihood),
  BIC = sapply(aft_comb_results, function(x) x$BIC)
)

head(aft_model_results_df)
```












Attempt to fit weibul for 2 categorical variables
```{r}

wb_fit1 <- survreg(survival_object ~ Dvo_Code , data=data_excluded_days, dist="weibull")
wb_fit1
wb_fit2 <- survreg(survival_object ~ Dvo_Code + Abbatoir, data=data_excluded_days, dist="weibull")
wb_fit2
wb_fit3 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed, data=data_excluded_days, dist="weibull")
wb_fit3
wb_fit4 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex, data=data_excluded_days, dist="weibull")
wb_fit4
wb_fit5 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm, data=data_excluded_days, dist="weibull")
wb_fit5
wb_fit6 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy, data=data_excluded_days, dist="weibull")
wb_fit6
wb_fit7 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy, data=data_excluded_days, dist="weibull")
wb_fit7
```

```{r}
wb_fit$
k=9
wb_fit1$loglik
AIC(wb_fit1, k=k)
akaike1 <- 2*k - 2*(wb_fit1$loglik[2])
akaike1
k=2
wb_fit2$loglik
AIC(wb_fit1, k=k)
akaike2 <- 2*k - 2*(wb_fit2$loglik[2])
akaike2
k=3
wb_fit3$loglik
AIC(wb_fit1, k=k)
akaike3 <- 2*k - 2*(wb_fit3$loglik[2])
akaike3
k=4
wb_fit4$loglik
AIC(wb_fit1, k=k)
akaike4 <- 2*k - 2*(wb_fit4$loglik[2])
akaike4
k=5
wb_fit5$loglik
AIC(wb_fit1, k=k)
akaike5 <- 2*k - 2*(wb_fit5$loglik[2])
akaike5
k=6
wb_fit6$loglik
AIC(wb_fit1, k=k)
akaike6 <- 2*k - 2*(wb_fit6$loglik[2])
akaike6

```

```{r}
wb_fit7 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy + T_2M, data=data_excluded_days, dist="weibull")
wb_fit7
wb_fit8 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy + T_2M + Total_Conditions_per_Animal, data=data_excluded_days, dist="weibull")
wb_fit8
wb_fit9 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy + T_2M + Total_Conditions_per_Animal + RH2M, data=data_excluded_days, dist="weibull")
wb_fit9
```

```{r}
wb_fit10 <- survreg(survival_object ~ Dvo_Code + Abbatoir + Breed + Sex + Abattoir_._Died_on_Farm + Beef_or_Dairy + T_2M + Total_Conditions_per_Animal + RH2M +No, data=data_excluded_days, dist="weibull")
wb_fit10
```

```{r}

```


```{r}
colnames(data_excluded_days)
```

create a function which will take all possible combinations of the variables
```{r}
library(combinat)
# create all possible combinations from variables
variables<-variable.names(data_excluded_days)
variables <- variables[-c(1, 3, 4, 5, 6, 10, 11, 15, 16, 17, 18)]
variables
 for (k in 1:length(variables)) {
    # Generate combinations of size k
    combs <- combn(variables, k, simplify = FALSE)
combs
```




delete rows were the days alive is 0
```{r}
excluded_days <- data$days_alive <=1
data_excluded_days <- data[!excluded_days, ]
summary(data_excluded_days$days_alive)

```

Fit a linear model in the Kaplan-Meier survival estimates after we transform them(log(-log)) to conform with the equations of Weibull distribution.
```{r}
baseline_survival_days <- survfit(Surv(days_alive, status) ~ 1, data = data_excluded_days)
baseline_survival_days
summary(baseline_survival_days$surv)
surv <- baseline_survival_days$surv
summary(surv)
exclude_surv0 <- as.numeric(which(surv<1e-12))
surv[exclude_surv0]
new_surv <- surv[-exclude_surv0]
summary(new_surv)
time <-baseline_survival_days$time 
summary(time)
new_time <- time[-exclude_surv0]
summary(new_time)


log_log_s <- log(-log(new_surv))
log_t <- log(new_time)

summary(log_log_s)
summary(log_t)

baseline_survival_days_lm <- lm(log_log_s ~ log_t)
baseline_survival_days_lm %>%
  broom::augment() %>%
  ggplot(aes(x = log_t)) +
  geom_point(aes(y = log_log_s)) +
  geom_line(aes(y = .fitted), linetype = 2, color = "goldenrod") +
  theme_light()
```


```{r}
coef(baseline_survival_days_lm)
```

fit distributions in dairy cattles
```{r}
train_dairy <- read.csv("./data/dairy_training_data.csv")
```

```{r}

excluded_months <- train_dairy$months_alive <=1
train_dairy_exclude_months<- train_dairy[!excluded_months, ]
summary(train_dairy_exclude_months$months_alive)

```


```{r}
validation_data <- read.csv("./data/dairy_validation_data.csv")
```




```{r}
variables <- variable.names(train_dairy)
variables <- variables[-c(2:5)]
variables
```

fit weibull for all categorical variables
```{r}
survival_object <- Surv(train_dairy_exclude_months$months_alive, train_dairy_exclude_months$status)
for (var_name in variables[2:68]) {
  cmd <- paste("wb_fit <- survreg(survival_object ~ ", var_name, ", data = train_dairy_exclude_months, dist = 'weibull')")
  eval(parse(text = cmd))
  # --ok
  
  cmd2 <- paste("new_dat <- expand.grid(", var_name, " = c('Other', '", var_name, "'),  survival = seq(.01, .99, by = .02) ) ")
  eval(parse(text = cmd2))
  
  cmd3 <- paste("new_dat <- new_dat %>% mutate( pred = map2(", var_name, 
    ", survival,  ~predict(wb_fit, type = 'quantile', p = 1 - .y, se = TRUE,  newdata = data.frame(", var_name, " = .x))),",
    "t = map_dbl(pred, ~pluck(.x, 'fit')),",
    "se = map_dbl(pred, ~pluck(.x, 'se.fit')),",
    "ucl = t + 1.96 * se,",
    "lcl = t - 1.96 * se)")
  
  eval(parse(text = cmd3))


  prcnt_lived_at_least <- function(x, isActiveVariable=1) {
    eval(parse(text=paste("num_active <- (validation_data %>% count(", var_name, "==isActiveVariable))[2, 2]")))
    eval(parse(text=paste("prcnt_lived_at_least_x <- (validation_data %>% count(months_alive >= x & ", var_name, "==isActiveVariable))[2, 2] / num_active")))
    return (prcnt_lived_at_least_x)
  }

 
  n_rows_new_dat <- nrow(new_dat)
  empirical_survival <- c()
  
  for (i in 1:n_rows_new_dat) {
    eval(parse(text=paste("isThisActive <- new_dat$", var_name, "[i]")))
    ti <- new_dat$t[i]
    empSurvi <- prcnt_lived_at_least(ti, isThisActive!="Other")
    empirical_survival <- append(empirical_survival, empSurvi)
  }
  
  new_dat$empirical_surv <- empirical_survival

  eval(parse(text=paste("my_plt_i <- new_dat %>%",
      "ggplot(aes(x = t, y=survival, color =", var_name, ")) + geom_line(linetype='dashed') + ",
      "geom_line(aes(y = empirical_surv, color = ", var_name, ")) +",
      "theme_light()")))
  print(my_plt_i)
  img_fname <- gsub(" ", "", paste("./images/weibull_fit/", var_name,".png", sep=""))
  ggsave(img_fname, plot = my_plt_i,  width = 8, height = 6)
}
```






fit exponential for all categorical variables

```{r}
survival_object <- Surv(train_dairy_exclude_months$months_alive, train_dairy_exclude_months$status)
for (var_name in variables[2:68]) {
  cmd <- paste("exp_fit <- survreg(survival_object ~ ", var_name, ", data = train_dairy_exclude_months, dist = 'exponential')")
  eval(parse(text = cmd))

  
  cmd2 <- paste("new_dat <- expand.grid(", var_name, " = c('Other', '", var_name, "'),  survival = seq(.01, .99, by = .02) ) ")
  eval(parse(text = cmd2))
  
  cmd3 <- paste("new_dat <- new_dat %>% mutate( pred = map2(", var_name, 
    ", survival,  ~predict(exp_fit, type = 'quantile', p = 1 - .y, se = TRUE,  newdata = data.frame(", var_name, " = .x))),",
    "t = map_dbl(pred, ~pluck(.x, 'fit')),",
    "se = map_dbl(pred, ~pluck(.x, 'se.fit')),",
    "ucl = t + 1.96 * se,",
    "lcl = t - 1.96 * se)")
  
  eval(parse(text = cmd3))


  prcnt_lived_at_least <- function(x, isActiveVariable=1) {
    eval(parse(text=paste("num_active <- (validation_data %>% count(", var_name, "==isActiveVariable))[2, 2]")))
    eval(parse(text=paste("prcnt_lived_at_least_x <- (validation_data %>% count(months_alive >= x & ", var_name, "==isActiveVariable))[2, 2] / num_active")))
    return (prcnt_lived_at_least_x)
  }

 
  n_rows_new_dat <- nrow(new_dat)
  empirical_survival <- c()
  
  for (i in 1:n_rows_new_dat) {
    eval(parse(text=paste("isThisActive <- new_dat$", var_name, "[i]")))
    ti <- new_dat$t[i]
    empSurvi <- prcnt_lived_at_least(ti, isThisActive!="Other")
    empirical_survival <- append(empirical_survival, empSurvi)
  }
  
  new_dat$empirical_surv <- empirical_survival

  eval(parse(text=paste("my_plt_i <- new_dat %>%",
      "ggplot(aes(x = t, y=survival, color =", var_name, ")) + geom_line(linetype='dashed') + ",
      "geom_line(aes(y = empirical_surv, color = ", var_name, ")) +",
      "theme_light()")))
  print(my_plt_i)
  img_fname <- gsub(" ", "", paste("./images/exponential_fit/", var_name,".png", sep=""))
  ggsave(img_fname, plot = my_plt_i,  width = 8, height = 6)
}
```















