---
title: "log_rank_tests"
author: "Marietta_Dalamanga"
date: "2024-07-14"
output:
  pdf_document: default
  html_document: default
---



```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(ggplot2)
library(scales)
library(survival)
library(survminer)
```



#### Log_rank test for beef_data


```{r}
beef_data <- read.csv("./data/beef_data.csv")
dairy_data <- read.csv("./data/dairy_data.csv")
```


We will perform log rank tests for all variables for survival in months in beef data

```{r}
variables <- colnames(dairy_data)
variables
```

log-rank test for herd id
```{r}
log_rank_months_herdid <- survdiff(Surv(time = beef_data$months_alive, event = beef_data$status) ~ as.factor(beef_data$Herd_Id), data=beef_data)
print(log_rank_months_herdid)
```

logrank test for temperature
```{r}
log_rank_months_temperature <- survdiff(Surv(time = beef_data$months_alive, event = beef_data$status) ~ beef_data$T_2M, data=beef_data)
print(log_rank_months_temperature)
```


log-rank test for each condition
```{r}
variables <- colnames(beef_data)

for(i in 14:41){
  variable_name <- variables[i]
  log_rank_months <- survdiff(Surv(time = beef_data$months_alive, event = beef_data$status) ~ as.factor(beef_data[[variable_name]]), data=beef_data)
  print(log_rank_months)
}
```


DVO, Herd id, Abattoir, Breed, Sex, place of death, total conditions per animal, temperature, RH2M
```{r}
for(i in c(1,4,5,6,7,8,9,12,13)){
  variable_name <- variables[i]
  log_rank_months <- survdiff(Surv(time = beef_data$months_alive, event = beef_data$status) ~ as.factor(beef_data[[variable_name]]), data=beef_data)
  print(log_rank_months)
}
```

further investigations for total conditions
first check if having no conditions is significant
```{r}
total_conditions_no_vs_all_other <- beef_data %>%
  mutate(conditions = ifelse(Total_Conditions_per_Animal == 0, 0, 1))
total_conditions_no_vs_all_other

log_rank_months_no_conditions_vs_other <- survdiff(Surv(time = total_conditions_no_vs_all_other$months_alive, event = total_conditions_no_vs_all_other$status) ~ as.factor(total_conditions_no_vs_all_other$conditions), data=total_conditions_no_vs_all_other)
print(log_rank_months_no_conditions_vs_other)
```


```{r}
total_conditions_more_than_9_vs_all_other <- beef_data %>%
  mutate(conditions = ifelse(Total_Conditions_per_Animal ==7, 0, 1))


log_rank_months_total_conditions_more_than_9_vs_all_other <- survdiff(Surv(time = total_conditions_more_than_9_vs_all_other$months_alive, event = total_conditions_more_than_9_vs_all_other$status) ~ as.factor(total_conditions_more_than_9_vs_all_other$conditions), data=total_conditions_more_than_9_vs_all_other)
print(log_rank_months_total_conditions_more_than_9_vs_all_other)
```



#### Log_rank test for dairy_data


We will perform log rank tests for all variables for survival in months

```{r}
variables <- colnames(dairy_data)
variables
```

log-rank test for each condition
```{r}
variables <- colnames(dairy_data)

for(i in 72){
  variable_name <- variables[i]
  print(variable_name)
  log_rank_months <- survdiff(Surv(time = dairy_data$months_alive, event = dairy_data$status) ~ as.factor(dairy_data[[variable_name]]), data=dairy_data)
  print(log_rank_months)
}
```


DVO, Herd id, Abattoir, Breed, Sex, place of death, total conditions per animal, temperature, RH2M
```{r}
for(i in c(1,4,5,6,7,8,9,12,13)){
  variable_name <- variables[i]
  log_rank_months <- survdiff(Surv(time = dairy_data$months_alive, event = dairy_data$status) ~ as.factor(dairy_data[[variable_name]]), data=dairy_data)
  print(log_rank_months)
}
```

further investigations for total conditions
first check if having no conditions is significant
```{r}
total_conditions_no_vs_all_other <- dairy_data %>%
  mutate(conditions = ifelse(Total_Conditions_per_Animal == 0, 0, 1))
total_conditions_no_vs_all_other

log_rank_months_no_conditions_vs_other <- survdiff(Surv(time = total_conditions_no_vs_all_other$months_alive, event = total_conditions_no_vs_all_other$status) ~ as.factor(total_conditions_no_vs_all_other$conditions), data=total_conditions_no_vs_all_other)
print(log_rank_months_no_conditions_vs_other)
```


Test is not significant only for 8 & 9 conditions but we have only 11 observations for 8 conditions and 2 observations for 9
```{r}
total_conditions_more_than_9_vs_all_other <- dairy_data %>%
  mutate(conditions = ifelse(Total_Conditions_per_Animal ==9, 0, 1))


log_rank_months_total_conditions_more_than_9_vs_all_other <- survdiff(Surv(time = total_conditions_more_than_9_vs_all_other$months_alive, event = total_conditions_more_than_9_vs_all_other$status) ~ as.factor(total_conditions_more_than_9_vs_all_other$conditions), data=total_conditions_more_than_9_vs_all_other)
print(log_rank_months_total_conditions_more_than_9_vs_all_other)
```
