---
title: "prepare_data_for_modelling"
author: "Marietta_Dalamanga"
date: "2024-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(ggplot2)
library(scales)
```

```{r}
memory.limit(size=70000)
```

```{r}
survival_data_without_outliers <- read.csv("./data/new_combined_data_without_outliers.csv")
```

I will drop the rows with na for now until Emily tells me what to do with this values
```{r}
survival_data_without_outliers <-drop_na(survival_data_without_outliers)
```


# keep only conditions with more than 1000 animals
We will keep only the animals with common condition


```{r}
colnames(survival_data_without_outliers)
conditions_columns <- colnames(survival_data_without_outliers[22:93])
conditions_columns
conditions_sums <- colSums(survival_data_without_outliers[, conditions_columns])
conditions_sums
rare_conditions_columns <- c(conditions_columns[conditions_sums < 2000])
rare_conditions_columns

```
keep only those columns

```{r}
filtered_rare_conditions <- survival_data_without_outliers
filtered_rare_conditions[rare_conditions_columns] <- list(NULL)
filtered_rare_conditions
```

Delete those rows where the animal has only rare conditions
```{r}
rare_conditions_data <- filtered_rare_conditions %>%
  filter(Total_Conditions_per_Animal > 0 & rowSums(filtered_rare_conditions[22:45]) == 0) 

common_conditions_data <- filtered_rare_conditions %>%
  filter(Total_Conditions_per_Animal== 0 | rowSums(filtered_rare_conditions[22:45]) >0 ) 
  
```
I will delete 2693 entries where the animals don't have any of the common conditions ### this was on the data before the herd size and the deletion of missing values
Now it is 2690 entries

```{r}
common_conditions_data
```

# Now we will delete those rows that the animals belong to a rare breed 

```{r}
N <- 8
top_breeds <- common_conditions_data %>%
  dplyr::count(Breed, sort=TRUE) %>%
  top_n(N, n) %>%
  pull(Breed)

top_breeds
common_conditions_breed_data <- common_conditions_data %>%
  filter(Breed %in% top_breeds)

nrow(common_conditions_breed_data)/nrow(common_conditions_data)* 100
```
Old dataset Top 8 breeds account for the 91.50% of the animals
We will keep only those
171019 rows were removed this is the 8.5% of our common_conditions_data

In the new dataset 170094 rows were removed the proportion of total animals is the same 
8.44% is on the survival data before removing the missing values
```{r}
(nrow(common_conditions_data)-nrow(common_conditions_breed_data)) / total_animals_no_outliers *100
```

```{r}
common_conditions_breed_data
```

save dataset in a csv_file to use it for parametric distributions
```{r}
write.csv(common_conditions_breed_data, "./data/common_conditions_breed_data.csv", row.names = FALSE)
```





##### all code below was for one hot encoding and split dataset. It is not longer used  ####
delete column that I will not use in modelling
these are: first index column, year, month, animal_is, Dob, Dod, days_alive and years_alive
We will keep the animal Id until the end of our data preparation
```{r}
colnames(filtered_condition_breeds)
filtered_condition_breeds <- filtered_condition_breeds[, -c(1,3,4,6,10,11,16,18)]
filtered_condition_breeds
```



```{r}
hist(filtered_condition_breeds$T_2M)
summary(filtered_condition_breeds$T_2M)

```


```{r}
hist(filtered_condition_breeds$RH2M)
summary(filtered_condition_breeds$RH2M)
```



scale 0 to 1 the numeric variables: Total_conditions_per_animal, T_2M and RH2M
```{r}
colnames(filtered_condition_breeds)
scale_0_1<- function(x){
  z <- (x-min(x))/ (max(x)-min(x))
  return(z)
}

filtered_condition_breeds[11:12] <- apply(filtered_condition_breeds[11:12], 2, scale_0_1)
filtered_condition_breeds[8] <- apply(filtered_condition_breeds[8], 2, scale_0_1)
summary(filtered_condition_breeds)

```

One hot encoding for the categorical variables
Create dummy variables for all categorical
```{r}
# Dvo_Code
trsf <- filtered_condition_breeds %>%
  mutate(Dvo_Code_present = 1) %>%
  pivot_wider(names_from = Dvo_Code, values_from = Dvo_Code_present, values_fill = list(Dvo_Code_present = 0))

colnames(trsf)[48] <- "NtArds"
colnames(trsf)
```

Here we prepend the string "ab_" to all abattoir codes so that when we apply one hot encodings later we get legit variable names
```{r}
#Abbatoir
prepend_abattoir = function(x, output){
  current_ab <- x[2]
  return(gsub(" ", "", paste("ab_", str_trim(current_ab), sep = "")))
}

modified_values <- apply(trsf, 1, prepend_abattoir)
trsf <- cbind(trsf, AbattoirFixed = modified_values)
head(trsf)
```

```{r}
trsf <- trsf %>%
  mutate(abbatoir_present = 1) %>%
  pivot_wider(names_from = AbattoirFixed, values_from = abbatoir_present, values_fill = list(abbatoir_present = 0))
colnames(trsf)
```

```{r}
# Breed
trsf <- trsf %>%
  mutate(breed_present = 1) %>%
  pivot_wider(names_from = Breed, values_from = breed_present, values_fill = list(breed_present = 0))
# Sex
trsf <- trsf %>%
  mutate(sex_present = 1) %>%
  pivot_wider(names_from = Sex, values_from = sex_present, values_fill = list(sex_present = 0))
# Abattoir_.Died_on_Farm
trsf <- trsf %>%
  mutate(place_of_death_present = 1) %>%
  pivot_wider(names_from = Abattoir_._Died_on_Farm, values_from = place_of_death_present, values_fill = list(place_of_death_present = 0))

trsf

```


```{r}
colnames(trsf)
```


# separate dataset into beef and dairy
```{r}
beef_data <- trsf %>%
  filter(Beef_or_Dairy=="Beef")

dairy_data <- trsf %>%
  filter(Beef_or_Dairy == "Dairy")
```

```{r}
nrow(beef_data) + nrow(dairy_data)

```

We will delete the columns that are of no use. These are: animal_id, Beef_or_dairy
```{r}
colnames(beef_data)
beef_data <- beef_data[, -c(1,3)]

dairy_data <- dairy_data[, -c(1,3)]
```

```{r}
colnames(beef_data)
colnames(beef_data)[73] <- "died_on_farm" 
colnames(beef_data)[72] <- "died_on_abattoir" 
colnames(beef_data)
```

```{r}
colnames(dairy_data)[73] <- "died_on_farm" 
colnames(dairy_data)[72] <- "died_on_abattoir" 
```

```{r}
write.csv(beef_data, "./data/beef_data.csv", row.names = FALSE)
write.csv(dairy_data, "./data/dairy_data.csv", row.names = FALSE)
```


we will split each data set into 3 parts 60% training set, 20% validation set and 20% test set.
the analysis will be performed for months_alive

```{r}
head(beef_data)
head(dairy_data)
```


Split the beef dataset into training and test set
```{r}
set.seed(123)
indeces_intrain_beef <-createDataPartition(y=beef_data$months_alive, p=0.6, list=FALSE)

beef_training_data <- beef_data[indeces_intrain_beef, ]
beef_testing_data <-beef_data[-indeces_intrain_beef, ]

nrow(beef_training_data) + nrow(beef_testing_data)
print(paste("observations in beef training set:", nrow(beef_training_data)))
print(paste("observations in beef test set:", nrow(beef_testing_data)))
```

Split the dairy dataset into training and test set
```{r}

set.seed(123)
indeces_intrain_dairy <-createDataPartition(y=dairy_data$months_alive, p=0.6, list=FALSE)

dairy_training_data <- dairy_data[indeces_intrain_dairy, ]
dairy_testing_data <-dairy_data[-indeces_intrain_dairy, ]

nrow(dairy_training_data) + nrow(dairy_testing_data)
print(paste("observations in dairy training set:", nrow(dairy_training_data)))
print(paste("observations in dairy test set:", nrow(dairy_testing_data)))

```


Further split the beef test set into half for test and half for validation set

```{r}
set.seed(123)
beef_val_ind <-createDataPartition(y=beef_testing_data$months_alive, p=0.5, list=FALSE)
beef_test_data <- beef_testing_data[beef_val_ind, ]
beef_validation_data <-beef_testing_data[-beef_val_ind, ]

nrow(beef_test_data) + nrow(beef_validation_data)
print(paste("observations in validation set:", nrow(beef_validation_data)))
print(paste("observations in test set:", nrow(beef_test_data)))
```

Dairy
```{r}
set.seed(123)
dairy_val_ind <-createDataPartition(y=dairy_testing_data$months_alive, p=0.5, list=FALSE)
dairy_test_data <- dairy_testing_data[dairy_val_ind, ]
dairy_validation_data <-dairy_testing_data[-dairy_val_ind, ]

nrow(dairy_test_data) + nrow(dairy_validation_data)
print(paste("observations in validation set:", nrow(dairy_validation_data)))
print(paste("observations in test set:", nrow(dairy_test_data)))
```



check sets:
beef
```{r}
nrow(beef_training_data) + nrow(beef_validation_data) + nrow(beef_test_data)

```

dairy
```{r}
nrow(dairy_training_data) + nrow(dairy_validation_data) + nrow(dairy_test_data)

```

save datasets in csv files
```{r}
write.csv(beef_training_data, "./data/beef_training_data.csv", row.names = FALSE)
write.csv(beef_validation_data, "./data/beef_validation_data.csv", row.names = FALSE)
write.csv(beef_testing_data, "./data/beef_test_data.csv", row.names = FALSE)
```


```{r}
write.csv(dairy_training_data, "./data/dairy_training_data.csv", row.names = FALSE)
write.csv(dairy_validation_data, "./data/dairy_validation_data.csv", row.names = FALSE)
write.csv(dairy_testing_data, "./data/dairy_test_data.csv", row.names = FALSE)
```

