---
title: "prepare_data_for_modelling"
author: "Marietta_Dalamanga"
date: "2024-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(plyr)
library(dplyr)
library(tidyr)
library(caret)
library(ggplot2)
library(scales)
```

```{r}
memory.limit(size=70000)
```


```{r}
survival_data_without_outliers <- read.csv("./data/new_combined_data_without_outliers.csv")
```


I will drop the rows with na for now until Emily tells me what to do with this values
```{r}
survival_data_without_outliers <-drop_na(survival_data_without_outliers)
```


# keep only conditions with more than 1000 animals
We will keep only the animals with common condition


```{r}
colnames(survival_data_without_outliers)
conditions_columns <- colnames(survival_data_without_outliers[22:93])
conditions_columns
conditions_sums <- colSums(survival_data_without_outliers[, conditions_columns])
conditions_sums
rare_conditions_columns <- c(conditions_columns[conditions_sums < 2000])
rare_conditions_columns

```
keep only those columns

```{r}
filtered_rare_conditions <- survival_data_without_outliers
filtered_rare_conditions[rare_conditions_columns] <- list(NULL)
filtered_rare_conditions
```

Delete those rows where the animal has only rare conditions
```{r}
rare_conditions_data <- filtered_rare_conditions %>%
  filter(Total_Conditions_per_Animal > 0 & rowSums(filtered_rare_conditions[22:45]) == 0) 

common_conditions_data <- filtered_rare_conditions %>%
  filter(Total_Conditions_per_Animal== 0 | rowSums(filtered_rare_conditions[22:45]) >0 ) 
  
```
I will delete 2693 entries where the animals don't have any of the common conditions ### this was on the data before the herd size and the deletion of missing values
Now it is 2690 entries

```{r}
common_conditions_data
```

# Now we will delete those rows that the animals belong to a rare breed 

```{r}
N <- 8
top_breeds <- common_conditions_data %>%
  dplyr::count(Breed, sort=TRUE) %>%
  top_n(N, n) %>%
  pull(Breed)

top_breeds
common_conditions_breed_data <- common_conditions_data %>%
  filter(Breed %in% top_breeds)

nrow(common_conditions_breed_data)/nrow(common_conditions_data)* 100
```
Old dataset Top 8 breeds account for the 91.50% of the animals
We will keep only those
171019 rows were removed this is the 8.5% of our common_conditions_data

In the new dataset 170094 rows were removed the proportion of total animals is the same 
8.44% is on the survival data before removing the missing values
```{r}
(nrow(common_conditions_data)-nrow(common_conditions_breed_data)) / total_animals_no_outliers *100
```

```{r}
common_conditions_breed_data
```

save dataset in a csv_file to use it for parametric distributions
```{r}
write.csv(common_conditions_breed_data, "./data/common_conditions_breed_data.csv", row.names = FALSE)
```



### split the data set to overcome problems with AFT NULL coefficients(overdetermined models) and Cox PH assumption
```{r}
survival_data_without_outliers <- read.csv("./data/common_conditions_breed_data.csv")

```


```{r}
summary(survival_data_without_outliers$days_alive)
```


create data partition

```{r}
set.seed(123)
split_indeces <- createDataPartition(survival_data_without_outliers$days_alive, p=0.005, list=FALSE)
split_data <- survival_data_without_outliers[split_indeces, ]
summary(split_data$days_alive)
big_data <- survival_data_without_outliers[-split_indeces, ]
```


```{r}

summary(big_data$days_alive)
summary(split_data$days_alive)
```

```{r}

table(big_data$Dvo_Code)
prop.table(table(big_data$Beef_or_Dairy))


```

change split data total conditions column to 0,1, 2 and >=3 conditions
```{r}
split_data$Total_Conditions_per_Animal[split_data$Total_Conditions_per_Animal >= 3] <- "3_or_more"
split_data$Total_Conditions_per_Animal
```

```{r}
write.csv(split_data, "./data/split_data.csv", row.names = FALSE)
```

change survival_data_without_outliers data total conditions column to 0,1, 2 and >=3 conditions
```{r}
survival_data_without_outliers$Total_Conditions_per_Animal[survival_data_without_outliers$Total_Conditions_per_Animal >= 3] <- "3_or_more"

```

change the name of the very long variables
```{r}
names(survival_data_without_outliers) <- c("Herd_Id", "Year", "Dvo_Code", "Month", "Animal_Id", "Abattoir", "Breed", "Sex", "Dob", "Dod", "place_of_death", "prod_type", "num_cond", "status",
                                           "days_alive", "months_alive", "years_alive", "T_2M", "RH2M", "herd_size", "No", "PLEUR.PNEUMONIA.LOCAL", "OEDEMA", "NEPHRITIS", "NAD", "CONTAMINATION",
                                           "TB.GENERALISED", "TB", "ABSCESS.PYAEMIA", "FLUKE.DAMAGE", "HYDRONEPHROSIS", "PLEUR.PNEUMONIA.GENERAL", "PETECHAE.HAEMORRHAGES", "INJECTION.SITES", 
                                           "PERITONITIS", "PLEURISY", "FACTORY.DAMAGE", "PERICARDITIS", "BRUISING", "SCAR.TISSUE", "TELANGECTASIS", "RESIDUES", "ARTHRITIS", "FASCIOLIASIS", "ACTINO" )

survival_data_without_outliers
```

change the values of Abattoir
```{r}
survival_data_without_outliers$Abattoir <- as.factor(survival_data_without_outliers$Abattoir)
levels(survival_data_without_outliers$Abattoir)
survival_data_without_outliers$Abattoir <- revalue(survival_data_without_outliers$Abattoir, c("32d8a591f643abd1ff6a998b128403c8962a55d8" ="ab_1", "332a6a1d2843f367c160a1c7c3fb2a28f0f8909e" ="ab_2",
                                                                                              "432d3a852c6074754d640e8d1369a1c56b0d5805"="ab_3",  "4f71277e271f09349a4d6d5f6c05acf635652015"="ab_4",
                                                                                              "52d480596328cca7948b6f1f31e1f2b18590cf10" ="ab_5", "568a97a7e2a6685902dd553c276b1837e0ebf06b"="ab_6",
                                                                                              "6f7d51dd8564e05e279faa791f97894ef141a4db"="ab_7", "7a032bbcfa7447fb21edc6e6f8d7bab0bed1bb16"="ab_8",
                                                                                              "87fc1df58026b8a228f9c4afa002b2b7722a8f09"="ab_9", "8a60fb1b9b40ed0fdc5d3c061e9388972f3bdce4"="ab_10",
                                                                                              "9032d4760ad9a99eb868b5d854cbae1492401dda"="ab_11", "94003c4910d6c601a698a3258918fc822b88dd0d"="ab_12",
                                                                                              "ae2f3299f13dc08c27a4bd4586901243017aa1c0"="ab_13", "ba747dbffafe62fb8bea541fd72df6d8c7afcb2f"="ab_14",
                                                                                              "cdf0a11f70869c640bfe6c23935ff3da4acfb7b0"="ab_15", "f833f5d69610136a115182502764c803be97ee16"="ab_16"))
survival_data_without_outliers
```


prepare training test and validation sets for machine learning
```{r}
set.seed(123)
test_validation_indeces <- createDataPartition(survival_data_without_outliers$days_alive, p=0.4, list=FALSE)
train_set <- survival_data_without_outliers[-test_validation_indeces, ]
test_validation_set <- survival_data_without_outliers[test_validation_indeces, ]
test_indeces <- createDataPartition(test_validation_set$days_alive, p=0.5, list=FALSE)
test_set <- test_validation_set[test_indeces, ]
validation_set <- test_validation_set[-test_indeces, ]
```

check the distribution of the response variable on three sets
```{r}
summary(survival_data_without_outliers$days_alive)
summary(train_set$days_alive)
summary(test_set$days_alive)
summary(validation_set$days_alive)
```

```{r}
paste("all observations are:", nrow(survival_data_without_outliers))
paste("train st has", nrow(train_set), "observations")
paste("test set has", nrow(test_set), "observations")
paste("validation set has", nrow(validation_set), "observations")
paste("sum of all three sets observations:", nrow(train_set)+nrow(test_set)+nrow(validation_set))
```

check the proportion of different variables on each set
```{r}
#Beef_or_Dairy
prop.table(table(survival_data_without_outliers$Beef_or_Dairy))
prop.table(table(train_set$Beef_or_Dairy))
prop.table(table(test_set$Beef_or_Dairy))
prop.table(table(validation_set$Beef_or_Dairy))
#DVO
prop.table(table(survival_data_without_outliers$Dvo_Code))
prop.table(table(train_set$Dvo_Code))
prop.table(table(test_set$Dvo_Code))
prop.table(table(validation_set$Dvo_Code))
#Breed
prop.table(table(survival_data_without_outliers$Breed))
prop.table(table(train_set$Breed))
prop.table(table(test_set$Breed))
prop.table(table(validation_set$Breed))
#Place of death
prop.table(table(survival_data_without_outliers$Abattoir_._Died_on_Farm))
prop.table(table(train_set$Abattoir_._Died_on_Farm))
prop.table(table(test_set$Abattoir_._Died_on_Farm))
prop.table(table(validation_set$Abattoir_._Died_on_Farm))
#Sex
prop.table(table(survival_data_without_outliers$Sex))
prop.table(table(train_set$Sex))
prop.table(table(test_set$Sex))
prop.table(table(validation_set$Sex))
#Total_conditions_per_animal
prop.table(table(survival_data_without_outliers$Total_Conditions_per_Animal))
prop.table(table(train_set$Total_Conditions_per_Animal))
prop.table(table(test_set$Total_Conditions_per_Animal))
prop.table(table(validation_set$Total_Conditions_per_Animal))
```
save sets on csv files
```{r}
write.csv(train_set, "./data/train.csv", row.names = FALSE)
write.csv(test_set, "./data/test.csv", row.names = FALSE)
write.csv(validation_set, "./data/validation.csv", row.names = FALSE)
```



